Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-12-19 Ben_Manuscripts/structure_paper/Draft.tex
revisions

  Nanostructured membrane materials have become increasingly popular for
  aqueous separations applications such as desalination and biorefinement because
  they offer the ability to control membrane architecture at the atomic scale
  permitting the design of solute-specific separation membranes.
  \cite{humplik_nanostructured_2011} Most membrane-based aqueous separations of
  small molecules can be achieved using reverse osmosis (RO) or nanofiltration
  (NF) \cite{van_der_bruggen_review_2003}. While RO and NF have seen many
  advances in the past few decades, they are far from perfect separation
  technologies.

@@

  Current commercial RO membranes are thin film composite membranes with a
  porous support layer and an active layer made of a dense polymer matrix. The
  membranes are unstructured with tortuous and polydisperse diffusion pathways.
  Separation occurs based on differences in solubility and diffusivity of solutes
  within the polymer matrix \cite{wijmans_solution-diffusion_1995}. With
  optimization, one can exploit these differences to create a functional
  selective barrier. RO operation costs are suboptimal because high feed
  pressures are required in order to achieve a useful solvent flux.

  %BJC : I don't know that this paragraph fits.  
  %MRS3: not really. Maybe take a look at Osuji's recent review (2016, I think) that addresses applications of nanostructured membranes and uses.  Could rewrite the introduction to be more about the general possibilities of nanostructured membranes, not just desalination.  It should be an introduction to what you will talk about in the paper, not to everything related.
  Improved RO membranes can make desalination more sustainable. Only 0.5 \% of
  the world's water is fresh. Desalination is an important source of potable
  water, necessary in order to keep up with demand generated by a rapidly growing
  global population. Although RO is currently the cheapest form of desalination,
  the cost must be further decreased in order to meet demand. A typical seawater
  desalination process requires feed pressures in the range of 800 to 1000 psi
  \cite{fritzmann_state---art_2007}. By designing better membranes for
  desalination, we can achieve higher water fluxes and reduce feed pressure
  requirements which will decrease the cost of fresh water production.
  \cite{humplik_nanostructured_2011}.

@@

  NF was introduced as an intermediate between RO and ultrafiltration, having
  the ability to separate organic matter and salts on the order of one nanometer
  in size. Explicit pore pathways running through the thickness of the membrane
  permit high solvent fluxes at lower pressures than RO. NF membranes typically
  have a surface charge resulting from ionizable groups which affords separation
  of ions smaller than the pore radius by the mechanism of Donnan exclusion
  \cite{van_der_bruggen_review_2003}. NF is often used as a precursor to reverse
  osmosis since it can efficiently remove a significant portion of dissolved
  solutes. Following NF pretreatment, RO can further purify the permeate. 
  Unfortunately, NF membranes, like RO, possess a pore size
  distribution which limits their ability to perform precise separations
  \cite{bowen_modelling_2002}. An NF membrane with high selectivity could offer
  the versatility to separate ions as well as small charged and uncharged
  molecules to the same degree as RO and at a significantly lower cost.

%MRS3: I think that this paragraph probably should be couched in terms of what they COULD do, since are not widely used yet.
%MRS3: if talking about existing membranes, should cite performance.
  Nanostructured membranes can bypass many of the performance issues which
  plague traditional NF and RO membranes. One can accomplish targeted separations
  with high selectivity by tuning shape, size and functionality of the molecular
  building blocks which form these materials. As a result, solute rejecting pores
  can have their sizes tuned uniformly, resulting in strict size cut-offs.
  Entirely different mechanisms may govern transport in a given nanostructured
  material which can inspire novel separation techniques.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II} phase LLC membrane studies have been limited primarily to
  Na-GA3C11 with some characterization done after minor structural modifications.
  Resel et al. varied the length of the monomer tails and the counterion used and
  observed its affect on pore spacing \cite{resel_structural_2000}.  Their study
  offers no insight into dynamics within the pore. In a later study of rejection
  performance, it was shown that membranes formed by Na-GA3C11 
%can not 
  cannot perform
  separations of solutes less than 1.2 nm in diameter because the pores are too
  large \cite{zhou_supported_2005}.  We do not yet understand how to controllably
  reduce the effective pore size or how to tune the chemical environment in the
  nanopores for effective water desalination and small organic molecule
  separations. The only source of predictive modeling for LLC systems have been
  macroscopic models which likely do not adequately describe transport at these
  length scales \cite{hatakeyama_water_2011}. It will be challenging to
  efficiently narrow down the large design space in a laboratory setting without
  a robust model.

  A molecular level understanding of LLC membrane structure, enabled by
  molecular dynamics simulations, 
%MRS3: suggest can provide, since we think it likely but cannot prove.
%will 
  can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail (Fig. \ref{fig:detail}).
  We can directly observe solute transport and suggest governing mechanisms. We
  can observe how the choice of head group may influence pore size for size
  exclusion driven separations. We can interchange counterions which may
  influence both the pore size and the strength of the Donnan potential which
  affects the degree to which the membrane can exclude charged species.

  % BJC: TODO: Make new rendering of wedges
  %            Change color scheme 
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
%MRS3: tweaked
%  \caption{(a) Our previous understanding of the pore structure allows us to speculate
%	   about separation behavior. (b) A detailed molecular model will allow us to
%	   directly observe solute transport. Here, four stacked layers of 5 monomers
%           are pictured atomistically. The hydrophilic region is in red and the 
%	   hydrophobic region is colored blue.}~\label{fig:detail}
  \caption{Our previous level of understanding of the pore structure (a) only allows us to speculate
	   about transport behavior in the pore. A detailed molecular model (b) will allow us to
	   directly observe solute transport. Here, four stacked layers of 5 monomers
           are pictured from atomistic simulations. The hydrophilic region is in red and the 
	   hydrophobic region is colored blue.}~\label{fig:detail}
  \end{figure}

@@

  Despite having structural data, there is still information which experiment
  cannot definitively answer. There are several key questions that we will
  investigate.
% which will be laid out and numbered in subsequent paragraphs.

  %BJC: commenting out :There has been no definitive answer in literature
  %regarding the number of monomers in each layer.  MRS3: the intro to the list
  %talks about only ``numbers of monomers in each layer'', whereas the list
  %below talks about several questions relating to layers. Can you rephrase so
  %it's more general? Maybe just take out that sentence above.

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations versus a sandwiched
  conformation ~\cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
%MRS3: perhaps ``in this study, we compare'', since the rest of the paper does not take place 
in the the future of the author. Apply throughout the paper.
  \cite{waller_hybrid_2006,ringer_effect_2006}. We will compare simulated X-ray
  diffraction patterns to experiment in order to deduce which stacking 
  configurations is most likely.

@@

  There remains the possibility that there is more than one metastable state
  associated with a given LLC system. Simulating a membrane atomistically will
%MRS3: rephrase.
%  require many atoms 
requires large systems
which limits the timescales acessible with MD. It is
  reasonable to expect that we will generate configurations which are kinetically
  trapped in a metastable free energy basin. We must be able to identify which
  state is produced experimentally.

@@

  %BJC: Would it be better to describe each reflection as a numbered list (like with the questions above?)
  %MRS2: I like names better.
  We used experimental small-angle X-ray scattering (SAXS) data from
  %MRS3: note that ~ is a space, so you don't need to do Fig.~ ref.  The ~ keeps the Fig.~\ref together instead of being split over lines.
  \cite{feng_thin_2016} (Fig.~\ref{fig:SAXS}) and wide angle X-ray scattering
  (WAXS) data (Fig.~\ref{fig:WAXS}, produced as described in
  \cite{feng_scalable_2014}) for comparison to our model. We rely primarily on the 2D WAXS data
  since it encodes all structural details down to the sub-nm scale.  There are
  five major features of interest present in the 2D experimental pattern shown in
  Figure \ref{fig:WAXS}. 
%MRS3: break this into an enumerated list, rather than a wall of text.
The first is located at $q_z$ = 1.7 \angstrom$^{-1}$,
  corresponding to a real space separation of 3.7 \angstrom~.  The reflection is
  attributed to $\pi$-$\pi$ stacking between aromatic rings in the direction
  perpendicular to the membrane plane, or z-axis \cite{feng_scalable_2014}. For
  simplicity, this reflection will be referred to as R-$\pi$. A weak intensity
  line is located at exactly half the $q_z$ value of R-$\pi$ ($q_z$ = 0.85
  \angstrom$^{-1}$), corresponding to a real space periodic spacing of 7.4
  \angstrom~. This reflection has been interpreted as 2\textsubscript{1} helical
  ordering of aromatic rings along the z axis, meaning that if the positions of the
  aromatic rings can be traced by a helix, then for each full turn in the helix, one
  will encounter two aromatic rings. For this reason it will be referred to as
  R-helix. A third major reflection is marked by a low intensity ring located at
  r = 1.4 \angstrom$^{-1}$. The real space separation corresponds to 4.5 \angstrom~
  which is characteristic of the average spacing between packed alkane chains.
  This reflection will be called R-alkanes. Within R-alkanes, are four spots of
  higher relative intensity which will be called R-spots. All are located
  $\sim 37$ degrees from the $q_z$ axis in their respective quadrants. In many
  liquid crystal systems this can be explained by the tilt angle of the alkane
  chains with respect to the membrane plane. The final feature corresponds to the
  spacing and symmetry of the d\textsubscript{100} plane which can be related to
  the distance between pores.  The feature, which will be called R-pores, is
  characterized by dots along $q_z$ = 0. The spacing between dots is indicative
  of the hexagonal symmetry of the packed pores. The same information at higher
  resolution is obtained using a SAXS setup.

  %MRS3: this paragraph seems to be a little out of place: should go before a lot of the introductory material, since it lays out what you will be doing in the paper.

  In this study, we build a significantly more realistic atomistic model of LLC
  membranes than, to our knowledge, has ever previously been done, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the model using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure which have been drawn from X-ray diffraction (XRD) experiments
  and in matching ionic conductivity measurements \cite{feng_thin_2016}.

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
%MRS; would be good to have all feature labels agree: either can edit to put the names on, or have a key relating numbers to the names you use.
	\caption{(a) The repeat spacing in the 1D small angle X-ray scattering pattern
		is characteristic of hexagonal packing. The leading peak represents the
		distance between the d\textsubscript{100} planes. Using this distance, we know
		that the distance between pore centers is 4.12 nm. (b) 2D wide angle X-ray
		scattering gives details about repeating features less than 1 nanometer
		apart. Experimentalists have justified each of the 5 major reflections
		present as follows: (I) Aromatic head group $\pi-\pi$ stack 3.7 \AA~apart.
		(II) Monomers arrange vertically in a 2\textsubscript{1} helix. (III) Alkane
 		chain tails pack 4.5 \AA~apart. (IV) Monomer tails tilt with respect to the 
 		membrane plane. (V) As derived from SAXS, pores pack hexagonally and are 
		spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

@@

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a parameterized
  monomer. Monomers were cooled from 1000K to 50K over 10 nanoseconds.  A low
%MRS3: need to describe this in more detail, but can be in the supplementary. 
  energy configuration was randomly pulled from the trajectory and charges were
  reassigned using molcharge.
%MRS3: need to say what molcharge is.
  Using the new charges, the monomer system was
  annealed again and a random monomer configuration was pulled from the
  trajectory to be used for full system construction (Figure~\ref{fig:python}a).

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of twenty stacked monomer layers with periodic
  continuity in the z direction, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. A small number of layers is
  preferred in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomer
  layers in each pore in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. %This point will be explained in more detail later.
%MRS3: isn't there some observations you need at least 10?  
\subsection{Monomer Placement}

@@

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in the
  NVT ensemble which allows the monomer tails to settle without disrupting the
  ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, the restraints are released linearly 
%MRS3: be more precise? you could reduce it from 10 in steps of 0.000001 kJ/mol, and that would be linear.
%until there is no more
%  restraining potential. 
until the restraining potential is zero.
The resulting unrestrained structure is allowed to
  equilibrate for 5 ns in the NPT ensemble with pressure controlled by the
  berendsen barostat. Next, we run long NPT equilibration simulations for at
  least 400 ns using the Parrinello-Rahman barostat with a time constant of 10 ps.

@@

  In order to fully match synthetic procedures, we created a crosslinking
  algorithm that can be applied to equilibrated structures. The purpose of 
  crosslinking
  is to maintain macroscopic alignment of the crystalline domains, 
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. The
  algorithm was developed based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, the initiator was simulated as hydrogen and made
  present in the simulation by including them in all possible locations where an 
  addition could occur as dummy atoms. The crosslinking procedure is carried out
  iteratively. During each iteration, bonding carbon atoms are chosen based on
  a distance cut-off. The topology is updated with new bonds and dummy hydrogen
  atoms are changed to appropriate hydrogen types. Head-to-tail addition was the
  only propagation mode considered due to its dominance in the real system.
%MRS3: citation why it dominates the real system. 
  Direction of attack was not considered because the resultant mixture is
  racemic.

  Our implementation requires long simulation times 
%MRS3: specify long.  2 days?  2 months?
to achieve high degrees
  of crosslinking. For that reason we did not crosslink all systems tested, but only the most
  promising structure. We show that crosslinking does not significantly 
  change any of our drawn conclusions in Section 3.6.

@@

  Using equilibrated structures, we carry out various calculations to
  characterize the system. We define the point at which a system is
  equilibrated based on when the distance between pores stops changing.
%MRS: how did you determine when it stopped changing?  What statistical tests? By eye?

@@

%  \subsection{Determination of Nanoscopic Structural Details}
  \subsection{Determining the Spatial Configuration of Monomers}

@@

%MRS3: specify which ones are referred to by neither.
  Neither assembly deviates from its initial head group arrangement. R-helix is still 
  faintly visible in the parallel displaced configuration and is absent in the 
  sandwiched simulated diffraction pattern. The spectroscopic signatures are unique to
  the two different head group configurations.

>>>>

  Nanostructured membrane materials have become increasingly popular for
  aqueous separations applications such as desalination and biorefinement because
  they offer the ability to control membrane architecture at the atomic scale
  permitting the design of solute-specific separation membranes
  \cite{humplik_nanostructured_2011}. Most membrane-based aqueous separations of
  small molecules can be achieved using reverse osmosis (RO) or nanofiltration
  (NF) \cite{van_der_bruggen_review_2003}. While RO and NF have seen many
  advances in the past few decades, they are far from perfect separation
  technologies.

  Current commercial RO and NF membranes suffer from inherent limitations of
  their fabrication. Although scalable, each process has a degree of
  stochasticity which makes overcoming the permeability-selectivity tradeoff a
  challenge. Namely, it is difficult to increase the permeability of a desired
  species, while maintaining the same retention of an undesired species.

  Commercial RO membranes are thin film composite membranes with a porous
  support layer and an active layer made of a polymer matrix formed by
  interfacial polymerization of a diamine. During the membrane fabrication,
  aqueous diamine solution is allowed to penetrate into the polysulfone support
  layer which is then immersed in trimesoyl chloride, a compound that is
  immiscible in water. Diamine polymerizes with trimesoyl chloride at the
  solution interface and creates a thin film that physically adheres to the
  polysulfone support. The polymer matrix is dense and highly crosslinked
  which leads to low permeability.

  NF membranes are typically made by a phase inversion process. During phase
  inversion, a thin polymer film is precipitated out of a polymer solution.
  Through the use of a non-solvent, a polymer-rich and polymer-poor phase is
  formed. A membrane is created as the polymer-poor phase forms pores in the
  polymer-rich phase. Resultant pores are polydisperse in size which hurts
  membrane selectivity. A second technique used to create NF membranes is call
  track etching in which a polymer film is bombarded with charged particles, then
  chemically etched to create pores. The pores are uniform which benefits
  selectivity, however the membrane has a low porosity and subsequent low
  permeability.

@@

%  Current commercial RO membranes are thin film composite membranes with a
%  porous support layer and an active layer made of a dense polymer matrix. The
%  membranes are unstructured with tortuous and polydisperse diffusion pathways.
%  Separation occurs based on differences in solubility and diffusivity of solutes
%  within the polymer matrix \cite{wijmans_solution-diffusion_1995}. With
%  optimization, one can exploit these differences to create a functional
%  selective barrier. RO operation costs are suboptimal because high feed
%  pressures are required in order to achieve a useful solvent flux.

  %BJC : I don't know that this paragraph fits.  
  %MRS3: not really. Maybe take a look at Osuji's recent review (2016, I think) that addresses applications of nanostructured membranes and uses.  Could rewrite the introduction to be more about the general possibilities of nanostructured membranes, not just desalination.  It should be an introduction to what you will talk about in the paper, not to everything related.
 % Improved RO membranes can make desalination more sustainable. Only 0.5 \% of
 % the world's water is fresh. Desalination is an important source of potable
 % water, necessary in order to keep up with demand generated by a rapidly growing
 % global population. Although RO is currently the cheapest form of desalination,
 % the cost must be further decreased in order to meet demand. A typical seawater
 % desalination process requires feed pressures in the range of 800 to 1000 psi
 % \cite{fritzmann_state---art_2007}. By designing better membranes for
 % desalination, we can achieve higher water fluxes and reduce feed pressure
 % requirements which will decrease the cost of fresh water production.
 % \cite{humplik_nanostructured_2011}.

@@

  %NF was introduced as an intermediate between RO and ultrafiltration, having
  %the ability to separate organic matter and salts on the order of one nanometer
  %in size. Explicit pore pathways running through the thickness of the membrane
  %permit high solvent fluxes at lower pressures than RO. NF membranes typically
  %have a surface charge resulting from ionizable groups which affords separation
  %of ions smaller than the pore radius by the mechanism of Donnan exclusion
  %\cite{van_der_bruggen_review_2003}. NF is often used as a precursor to reverse
  %osmosis since it can efficiently remove a significant portion of dissolved
  %solutes. Following NF pretreatment, RO can further purify the permeate. 
  %Unfortunately, NF membranes, like RO, possess a pore size
  %distribution which limits their ability to perform precise separations
  %\cite{bowen_modelling_2002}.

  The permeability-selectivity tradeoff has the potential to be overcome by
  designing membranes at the molecular level. Next-geneartion nanostructured
  membranes with high selectivity permitted by a precisely controlled pore size
  and high permeability allowed by its porous architecture, have the potential to
  replace traditional RO and NF membrane technologies.

%MRS3: I think that this paragraph probably should be couched in terms of what they COULD do, since are not widely used yet.
%MRS3: if talking about existing membranes, should cite performance.
%  Nanostructured membranes can bypass many of the performance issues which
%  plague traditional NF and RO membranes. One can accomplish targeted separations
%  with high selectivity by tuning shape, size and functionality of the molecular
%  building blocks which form these materials. As a result, solute rejecting pores
%  can have their sizes tuned uniformly, resulting in strict size cut-offs.
%  Entirely different mechanisms may govern transport in a given nanostructured
%  material which can inspire novel separation techniques.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II} phase LLC membrane studies have been limited primarily to
  Na-GA3C11 with some characterization done after minor structural modifications.
  Resel et al. varied the length of the monomer tails and the counterion used and
  observed its affect on pore spacing \cite{resel_structural_2000}.  Their study
  offers no insight into dynamics within the pore. In a later study of rejection
  performance, it was shown that membranes formed by Na-GA3C11 cannot perform
  separations of solutes less than 1.2 nm in diameter because the pores are too
  large \cite{zhou_supported_2005}.  We do not yet understand how to controllably
  reduce the effective pore size or how to tune the chemical environment in the
  nanopores for effective water desalination and small organic molecule
  separations. The only source of predictive modeling for LLC systems have been
  macroscopic models which likely do not adequately describe transport at these
  length scales \cite{hatakeyama_water_2011}. It will be challenging to
  efficiently narrow down the large design space in a laboratory setting without
  a robust model.

  %MRS3: this paragraph seems to be a little out of place: should go before a lot of the introductory material, since it lays out what you will be doing in the paper.
  %BJC2: Paragraph moved up here from well below.
  In this study, we build a significantly more realistic atomistic model of LLC
  membranes than, to our knowledge, has ever previously been done, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the model using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure which have been drawn from X-ray diffraction (XRD) experiments
  and in matching ionic conductivity measurements \cite{feng_thin_2016}.

  A molecular level understanding of LLC membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}).  We can directly observe solute transport and suggest
  governing mechanisms. We can observe how the choice of head group may influence
  pore size for size exclusion driven separations. We can interchange counterions
  which may influence both the pore size and the strength of the Donnan potential
  which affects the degree to which the membrane can exclude charged species.

  % BJC: TODO: Make new rendering of wedges
  %            Change color scheme 
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
  \caption{Our previous level of understanding of the pore structure (a) only allows us to speculate
	   about transport behavior in the pore. A detailed molecular model (b) will allow us to
	   directly observe solute transport. Here, four stacked layers of 5 monomers
           are pictured from atomistic simulations. The hydrophilic region is in red and the 
	   hydrophobic region is colored blue.}~\label{fig:detail}
  \end{figure}

@@

  Despite having structural data, there is still information which experiment
  cannot definitively answer. There are several key questions that we will
  investigate.

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations versus a sandwiched
  conformation ~\cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
  \cite{waller_hybrid_2006,ringer_effect_2006}. In this study, we compare
  simulated X-ray diffraction patterns to experiment in order to deduce which
  stacking configurations is most likely.

@@

  There remains the possibility that there is more than one metastable state
  associated with a given LLC system. Simulating a membrane atomistically
  requires many atoms which limits the timescales acessible with MD. It is
  reasonable to expect that we will generate configurations which are kinetically
  trapped in a metastable free energy basin. We must be able to identify which
  state is produced experimentally.

@@

  We used experimental small-angle X-ray scattering (SAXS) data from
  \cite{feng_thin_2016} (Fig.~\ref{fig:SAXS}) and wide angle X-ray scattering
  (WAXS) data (Fig.~\ref{fig:WAXS}, produced as described in
  \cite{feng_scalable_2014}) for comparison to our model. We rely primarily on the 2D WAXS data
  since it encodes all structural details down to the sub-nm scale.  There are
  five major features of interest present in the 2D experimental pattern shown in
  Figure \ref{fig:WAXS}.

  	\item The first is located at $q_z$ = 1.7 \angstrom$^{-1}$, corresponding to
	a real space separation of 3.7 \angstrom~.  The reflection is attributed to
	$\pi$-$\pi$ stacking between aromatic rings in the direction perpendicular to
	the membrane plane, or z-axis \cite{feng_scalable_2014}. For simplicity, this
	reflection will be referred to as R-$\pi$.

  	\item A weak intensity line is located at exactly half the $q_z$ value of
	R-$\pi$ ($q_z$ = 0.85 \angstrom$^{-1}$), corresponding to a real space periodic
	spacing of 7.4 \angstrom~. This reflection has been interpreted as
	2\textsubscript{1} helical ordering of aromatic rings along the z axis, meaning
	that if the positions of the aromatic rings can be traced by a helix, then for
	each full turn in the helix, one will encounter two aromatic rings. For this
	reason it will be referred to as R-helix.

	\item A third major reflection is marked by a low intensity ring
	located at r = 1.4 \angstrom$^{-1}$. The real space separation corresponds to
	4.5 \angstrom~ which is characteristic of the average spacing between packed
	alkane chains.  This reflection will be called R-alkanes.

	\item Within R-alkanes, are four spots of higher relative intensity
	which will be called R-spots. All are located $\sim 37$ degrees from the $q_z$
	axis in their respective quadrants. In many liquid crystal systems this can be
	explained by the tilt angle of the alkane chains with respect to the membrane
	plane.

	\item The final feature corresponds to the spacing and symmetry of the
	d\textsubscript{100} plane which can be related to the distance between pores.
	The feature, which will be called R-pores, is characterized by dots along $q_z$
	= 0. The spacing between dots is indicative of the hexagonal symmetry of the
	packed pores. The same information at higher resolution is obtained using SAXS.

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
%MRS3: would be good to have all feature labels agree: either can edit to put the names on, or have a key relating numbers to the names you use.
%BJC2: I will annotate it with the names R-pi, R-alkanes etc. Is that what you mean?
	\caption{(a) The repeat spacing in the 1D small angle X-ray scattering pattern
		is characteristic of hexagonal packing. The leading peak represents the
		distance between the d\textsubscript{100} planes. Using this distance, we know
		that the distance between pore centers is 4.12 nm. (b) 2D wide angle X-ray
		scattering gives details about repeating features less than 1 nanometer
		apart. Experimentalists have justified each of the 5 major reflections
		present as follows: (I) Aromatic head group $\pi-\pi$ stack 3.7 \AA~apart.
		(II) Monomers arrange vertically in a 2\textsubscript{1} helix. (III) Alkane
 		chain tails pack 4.5 \AA~apart. (IV) Monomer tails tilt with respect to the 
 		membrane plane. (V) As derived from SAXS, pores pack hexagonally and are 
		spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

@@

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a parameterized
  monomer. Monomers were cooled from 1000K to 50K over 10 nanoseconds.  A low
%MRS3: need to describe this in more detail, but can be in the supplementary. 
%BJC2: would it be best to include the parameterization script
  energy configuration was randomly pulled from the trajectory and charges were
  reassigned using molcharge.
%MRS3: need to say what molcharge is.
%BJC2: I say in paragraph above : "Atomic charges were assigned using the am1bccsym method of 
%molcharge shipped with QUACPAC from ...". Is that not enough?
  Using the new charges, the monomer system was
  annealed again and a random monomer configuration was pulled from the
  trajectory to be used for full system construction (Figure~\ref{fig:python}a).
  See supplemental information for further detail.

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of twenty stacked monomer layers with periodic
  continuity in the z direction, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. A small number of layers is
  preferred in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomer
  layers in each pore in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. %This point will be explained in more detail later.
%MRS3: isn't there some observations you need at least 10?  
%BJC2: When I first set up the system, there was a gap and a small number of layers 
% appeared to cause the phase to disassociate. However, I can get a pretty small
% number of layers in the infinite pore system, but the diffraction is low resolution.

\subsection{Monomer Placement}

@@

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in
  the NVT ensemble which allows the monomer tails to settle without disrupting
  the ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, the restraints are reduced in a sequence with values of
  8, 3, 2, 1, and 0 KJ mol$^{-1}$ nm$^{-2}$ respectively. The resulting
  unrestrained structure is allowed to equilibrate for 5 ns in the NPT ensemble
  with pressure controlled by the berendsen barostat. Next, we run long NPT
  equilibration simulations for at least 400 ns using the Parrinello-Rahman
  barostat with a time constant of 10 ps.

@@

  In order to fully match synthetic procedures, we created a crosslinking
  algorithm that can be applied to equilibrated structures. The purpose of 
  crosslinking
  is to maintain macroscopic alignment of the crystalline domains, 
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. The
  algorithm was developed based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, the initiator was simulated as hydrogen and made
  present in the simulation by including them in all possible locations where an 
  addition could occur as dummy atoms. The crosslinking procedure is carried out
  iteratively. During each iteration, bonding carbon atoms are chosen based on
  a distance cut-off. The topology is updated with new bonds and dummy hydrogen
  atoms are changed to appropriate hydrogen types. Head-to-tail addition was the
  only propagation mode considered due to its dominance in the real system.
%MRS3: citation why it dominates the real system.
%BJC2: Yea. There are a few citations that I need to make. I know they exist
% but I need to track them down. I've prioritized that last. 
  Direction of attack was not considered because the resultant mixture is
  racemic.

  Our implementation requires long simulation times to achieve high degrees of
  crosslinking. A typical crosslinking procedure can take up to 24 hours. In
  order to collect equilibrated data, further NPT simulation is necessary. We
  typically run a crosslinked system for an additional 100 ns to allow the system
  to readjust. For those reasons we did not crosslink all systems tested, but only
  the most promising structure. We show that crosslinking does not significantly
  change any of our drawn conclusions in Section 3.6.

@@

  Using equilibrated structures, we carry out various calculations to
  characterize the system. We define the point at which a system is
  equilibrated based on when the distance between pores stops changing.
  We determined when the distances stopped changing by applying the statistical
  test, \texttt{pymbar.timeseries.detectEquilibration}, to the time series. 
  Simulations of 400 ns give at least 50 ns of equilibrated simulation trajectory.
  %BJC2: Do I cite the github repository for pymbar?

@@

  \subsection{Determining the Spatial Configuration of Monomers}

@@

  The sandwiched and parallell displaced assemblies do not deviate from their
  initial head group arrangement. R-helix is still faintly visible in the
  parallel displaced configuration and is absent in the sandwiched simulated
  diffraction pattern. The spectroscopic signatures are unique to the two
  different head group configurations.
